Biomass predictions and VOCC

### 2019-08-13 analysis/VOCC/4-vocc-by-spp.Rmd
use log_change as response; save lmer coefs

Species
```{r global_options, include=FALSE}
# library(dplyr)
# library(ggplot2)
# library(gfplot)
# library(gfdata)
# library(sdmTMB)
# library(gfranges)
# 
# knitr::opts_chunk$set(
#   fig.width = 11, fig.height = 8.5,
#   echo = FALSE, warning = FALSE, message = FALSE
# )
```

@@

climate <- "do and temperature"
max_thresholds <- c(Inf, 0.75)
min_thresholds <- c(0.25, Inf)
critical_thresholds <- c(-0.25, 0.75)
optimal_values <- c(2, 6.5)
```

@@

# #   #climate <- "do and temperature"
# max_thresholds <- c(Inf, 0.5)
# min_thresholds <- c(0.1, Inf)
# # critical_thresholds <- c(-0.25, 0.75)
# optimal_values <- c(2, 6.5)

@@

if (length(vectors[ ,2]) == 0) {
  other_coefs <- readRDS("data/vocc-by-spp-sdmTMB-coefs.RDS")
  coefs <- other_coefs[1,]
  coefs$var <- "NA"
  coefs$beta <- "NA"
  coefs$se <- "NA"
  coefs$species <- spp 
  coefs$lifestage <- "NA"
  coefs$climate <- climate 
  coefs$min_thresholds <- min_string
  coefs$max_thresholds <- max_string
  coefs$optimum <- optimal_string
  coefs$start_time <- start_time
  coefs$timespan <- delta_t_total
  coefs$region <- region
  coefs$ssid_string <- ssid_string

  rbind(other_coefs, coefs)
  saveRDS(coefs, file = "data/vocc-by-spp-sdmTMB-coefs.RDS")

  stop("No vectors. Conditions saved to 'data/vocc-byspp-sdmTMB-coefs'.")
}

#glimpse(vectors)
```

@@

biomass_change <- make_trend_data(i,
  ssid = model_ssid,
  start_time = start_time,
  skip_time = skip_time,
  end_time = end_time,
  input_cell_size = input_cell_size,
  scale_fac = scale_fac,
  delta_t_total = delta_t_total,
  delta_t_step = delta_t_step,
  indices = indices,
  variable_names = "est_exp"
)

vectors <- vectors %>%
      mutate(
        start_cell = icell,
        vector_id = paste0(id, "_to_", tid),
        distance2 = distance*distance,
        source_end_value = e, # may need to turn into numeric vector later...
        target_value = target_values
      )

start <- biomass_change %>% 
  rename(start_density = est_exp_s, end_density = est_exp_e) %>% 
  select(icell, x, y, start_density, end_density)

@@

end <- biomass_change %>% 
  rename(start_density = est_exp_s, 
         end_density = est_exp_e, 
         target_X = x, target_Y = y) %>% 
  select(start_density, end_density, target_X, target_Y)

@@

source_density <- biomass_change %>% 
  rename(source_density = est_exp_s) %>% 
  select(source_density, x, y)

@@

  # calculate percent change in biomass
  mutate(perc_change = (end_density - start_density)/start_density)

@@

# calculate standardized density variables
vocc_by_sp <- vocc_by_sp %>% 
  mutate(
    std_source_density = (source_density - mean.density)/(sd.density*2), 
    mean_source_density = mean.density, sd_density = sd.density,
    mean_change = mean.change, sd_change = sd.change,
    std_change = (perc_change)/sd.change)

@@

vocc_by_sp <- mutate(vocc_by_sp,
  (!! mean.var) := icells[1, mean.var],
  (!! sd.var) := icells[1, sd.var],
  (!! std.var) := (UQ(rlang::sym(paste0(variable_names[j], "_e"))) - icells[1, mean.var])/ icells[1, sd.var],
  (!! std.var2) := ((UQ(rlang::sym(paste0(variable_names[j], "_e"))) - icells[1, mean.var])/ icells[1, sd.var])^2
  )
}

vocc_by_sp <- vocc_by_sp %>% 
  mutate(
    std_source_density = (source_density - mean.density)/(sd.density*2), 
    mean_source_density = mean.density, sd_density = sd.density,
    mean_change = mean.change, sd_change = sd.change,
    std_change = (perc_change)/sd.change)

@@

mature_plot <- lapply(variable_names, function(j) { 
ggplot(data, aes(x = UQ(rlang::sym(paste0("end_", j))), y = (end_density # - start_density
  ), colour = cell_type, #size = start_density, 
  group = cell_type)) +
    geom_point() +
  #geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
  geom_smooth(method="lm", formula = y ~ x + I(x^2) + I(x^3)) +
  #   , aes(x = end_climate, y = (end_density #- start_density
  #  )), inherit.aes = FALSE) +
  scale_size_continuous(guide = "none") +
  scale_y_continuous(trans = fourth_root_power) +
  ylab(paste("Final biomass")) +
  xlab(paste("Final value of", variable_names,"for all cells")) +
  scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
  gfplot::theme_pbs() + theme(legend.position = c(0.8, 0.8)) + 
  ggtitle(paste(survey, "mature", species))
})
print(mature_plot)
```

@@

mature_plot2 <- ggplot(data, aes(x = distance, y = std_change, 
  group = cell_type, size = start_density, colour = end_climate)) +
  geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
  geom_smooth(method="lm", formula = y~x+I(x^2) )+
  facet_wrap(~cell_type) +#, scales = "free") +
  scale_size_continuous(guide = "none") +
  # scale_y_continuous(trans = fourth_root_power) +
  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
  xlab(paste("Distance traveled to maintain <", min_thresholds, " decline or", max_thresholds, "increase in", climate,"")) +
  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", climate)) + 
  # scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
  gfplot::theme_pbs() + theme(legend.position = c(0.8, 0.8)) + 
  ggtitle(paste(survey, "mature", species))

mature_plot2
```

@@

mature_plot3 <- ggplot(data, aes(x = source_density, y = std_change, 
    colour = end_climate, 
    alpha = 1/(distance+ input_cell_size), #size = 1/distance,
    group = cell_type)) +
  geom_jitter(width = 0.01) +
  geom_smooth(method="lm") +
  facet_wrap(~cell_type) + #, scales = "free") +
  scale_alpha_continuous(guide = "none") +
  scale_size_continuous(guide = "none") +
  scale_x_continuous(trans = sqrt) +
  scale_y_continuous(trans = sqrt) +
  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
  xlab(paste("Source density (dark points are closer to source)")) +
  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", climate)) + 
  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
  ggtitle(paste(survey, "mature", species, "(mean biomass change =", round(data$mean_change[1], digits = 4), "kg/ha)"))

mature_plot3
```

@@

mature_plot4 <- ggplot(filter(data, distance < 10), aes(source_density, y = std_change , 
    colour = end_climate, 
    alpha = 1/(distance+ input_cell_size), #size = 1/distance,
    group = cell_type)) +
  geom_jitter(width = 0.01) +
  geom_smooth(method="lm", aes(group = cell_type)) +
  facet_wrap(~cell_type) +#, scales = "free") +
  scale_alpha_continuous(guide = "none") +
  scale_size_continuous(guide = "none") +
  scale_x_continuous(trans = sqrt) +
  #scale_y_continuous(trans = fourth_root_power) +
  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
  xlab(paste("Source density (dark points are closer to source)")) +
  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", climate)) + 
  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
  ggtitle(paste(survey, "mature", species, "(mean biomass change =", round(data$mean_change[1], digits = 4), "kg/ha)"))

mature_plot4
```

@@

mature_plot5 <- ggplot(data, aes(x = end_climate - initial_value , y = #std_change
    (end_density - start_density)
  , colour = cell_type, size = start_density)) +
  geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
  scale_size_continuous(guide = "none") +
  #scale_y_continuous(trans = fourth_root_power) +
  ylab(paste("kg/ha change in biomass over", vectors$timespan[1], "years")) +
  xlab(paste("Change in", climate,"")) +
  scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
  ggtitle(paste(survey, "mature", species))
mature_plot5
```

@@

```{r}
# test <- lmerTest::lmer(log(end_density) ~ cell_type * std_source_density + cell_type * distance + 
#             std_end_climate + std_end_climate2 + 
#             (1 | vector_id), 
#  #   offset = log(start_density), 
#  #   family = Gamma(link = "inverse")
#      data = data)
# hist(data$end_density)
# hist(data$std_change)
data <- vocc_by_sp[["adult"]]

test2 <- lmerTest::lmer(std_change ~ cell_type * std_source_density + 
    cell_type * distance + cell_type * distance2 +
    std_end_climate + std_end_climate2 + (1 | vector_id), REML = F, data = data)

summary(test2)
plot(test2)
```

@@

test3 <- robustlmm::rlmer(std_change ~ cell_type * std_source_density + 
    cell_type * distance + cell_type * distance2 + 
    std_end_climate + std_end_climate2 + (1 | vector_id), data = data)

@@

@@

m <- gbm::gbm(
  std_change ~ source_density + end_climate + distance + cell_type + X + Y , 
  data = data, 
  distribution = "gaussian",
  n.trees = 2000, 
  interaction.depth = 3, 
  shrinkage = 0.02
  )

@@

```{r}
plot(m,1)
plot(m,2)
plot(m,3)
plot(m,4)
plot(m,c(1,2))
plot(m,c(1,3))
plot(m,c(1,4))
plot(m,c(2,3))
plot(m,c(2,4))
plot(m,c(1,2,4))
plot(m,c(2,3,4))
plot(m,c(5,6))
```

@@

```{r}
data <- vocc_by_sp[["adult"]]
# Xmax <- quantile(data$X, 0.9995)
# Xmin <- quantile(data$X, 0.005)
# std_max <- quantile(data$std_change, 0.9995)
# 
# data <- data %>% filter ( X >= Xmin & X <= Xmax )

@@

```{r}
#data$end_climate2 <- data$end_climate^2
#data <- vocc_by_sp[["adult"]]
data$std_source_density2 <- data$std_source_density^2
test <- sdmTMB(data = data,
      std_change ~ cell_type * std_source_density + cell_type * std_source_density2 +
    cell_type * distance + cell_type * distance2 + 
      std_end_climate + std_end_climate2, 
      # std_change ~ cell_type * source_density + cell_type * distance  + end_climate + end_climate2,
      spde = spde, # control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = TRUE)
test
```

@@

Check final climate-biomass relationship
```{r}
idata <- vocc_by_sp[["imm"]]
#glimpse(data)

immature_plot1 <- ggplot(idata, aes(x = end_climate, y = (end_density # - start_density
  ), colour = cell_type, #size = start_density, 
  group = cell_type)) +
  geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
  geom_smooth(method="lm", formula = y ~ x + I(x^2) + I(x^3)) +
  #geom_smooth() +
  scale_size_continuous(guide = "none") +
  scale_y_continuous(trans = fourth_root_power) +
  ylab(paste("Final biomass")) +
  xlab(paste("Final value of", climate,"for all cells")) +
  scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
  gfplot::theme_pbs() + theme(legend.position = c(0.8, 0.8)) + 
  ggtitle(paste(survey, "immature", species))

immature_plot1
```

Check distance threshold of 50 looks reasonable
```{r warnings=FALSE}
idata <- vocc_by_sp[["imm"]]
#data <- filter(vocc_by_sp[["adult"]], distance > 1 & distance < 100)

immature_plot2 <- ggplot(idata, aes(distance, y = std_change, 
  group = cell_type, size = start_density, colour = end_climate)) +
  geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
  geom_smooth(method="lm", formula = y~x+I(x^2) )+
  facet_wrap(~cell_type) + #, scales = "free") +
  scale_size_continuous(guide = "none") +
  scale_y_continuous(trans = sqrt) +
  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
  xlab(paste("Distance traveled to maintain <", min_thresholds, " decline or", max_thresholds, "increase in", climate,"")) +
  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", climate)) + 
  # scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
  ggtitle(paste(survey, "immature", species))

immature_plot2
```

```{r warnings=FALSE}
idata <- vocc_by_sp[["imm"]]
#idata <- filter(vocc_by_sp[["imm"]], distance < 50)
#idata <- filter(vocc_by_sp[[""imm"]], distance > 4 & distance < 50)

immature_plot3 <- ggplot(idata, aes(x = source_density, y = std_change, 
    colour = end_climate, 
    alpha = 1/(distance+ input_cell_size), #size = 1/distance,
    group = cell_type)) +
  geom_jitter(width = 0.01) +
  geom_smooth(method="lm") +
  facet_wrap(~cell_type) + #, scales = "free") +
  scale_alpha_continuous(guide = "none") +
  scale_size_continuous(guide = "none") +
  scale_x_continuous(trans = sqrt) +
  scale_y_continuous(trans = sqrt) +
  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
  xlab(paste("Source density (dark points are closer to source)")) +
  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", climate)) + 
  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
  ggtitle(paste(survey, "immature", species, "(mean biomass change =", round(data$mean_change[1], digits = 4), "kg/ha)"))

immature_plot3
```

```{r warnings=FALSE}
idata <- vocc_by_sp[["imm"]]
idata <- filter(vocc_by_sp[["imm"]], distance < 50)

immature_plot4 <- ggplot(filter(idata, distance < 10), aes(source_density, y = std_change , 
    colour = end_climate, 
    alpha = 1/(distance+ input_cell_size), #size = 1/distance,
    group = cell_type)) +
  geom_jitter(width = 0.01) +
  geom_smooth(method="lm") +
  facet_wrap(~cell_type) +#, scales = "free") +
  scale_alpha_continuous(guide = "none") +
  scale_size_continuous(guide = "none") +
  scale_x_continuous(trans = sqrt) +
  scale_y_continuous(trans = sqrt) +
  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
  xlab(paste("Source density (dark points are closer to source)")) +
  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", climate)) + 
  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
  ggtitle(paste(survey, "immature", species, "(mean biomass change =", round(data$mean_change[1], digits = 4), "kg/ha)"))

immature_plot4
```

```{r}
idata <- vocc_by_sp[["imm"]]
idata <- filter(vocc_by_sp[["imm"]], distance < 50)

immature_plot5 <- ggplot(idata, aes(x = end_climate - initial_value , y = (end_density - start_density
  ), colour = cell_type, size = start_density)) +
  geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
  scale_size_continuous(guide = "none") +
  #scale_y_continuous(trans = fourth_root_power) +
  ylab(paste("kg/ha change in biomass over", vectors$timespan[1], "years")) +
  xlab(paste("Change in", climate,"")) +
  scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
  ggtitle(paste(survey, "immature", species))
immature_plot5
```

## Immature models
### Regular mixed-model

@@

```{r}
summary(mi)
```

>>>>

Species
```{r global_options, include=FALSE}
library(dplyr)
library(ggplot2)
# library(gfplot)
# library(gfdata)
library(sdmTMB)
library(gfranges)
# 
# knitr::opts_chunk$set(
#   fig.width = 11, fig.height = 8.5,
#   echo = FALSE, warning = FALSE, message = FALSE
# )
```

@@

climate <- "do and temperature"
max_thresholds <- c(Inf, 0.5)
min_thresholds <- c(0.1, Inf)
optimal_values <- c(2, 6.5)

# critical_thresholds <- c(-0.25, 0.75)

@@

# #   #climate <- "do and temperature"
# max_thresholds <- c(Inf, 0.5)
# min_thresholds <- c(0.1, Inf)
# optimal_values <- c(2, 6.5)

@@

if (length(vectors[ ,2]) == 0) {
  other_coefs <- readRDS(paste0("data/",climate_string,"-vocc-by-spp-sdmTMB-coefs.RDS"))
  coefs <- other_coefs[1,]
  coefs$var <- "NA"
  coefs$beta <- "NA"
  coefs$se <- "NA"
  coefs$model <- "NA"
  coefs$species <- spp 
  coefs$lifestage <- "NA"
  coefs$climate <- climate 
  coefs$min_thresholds <- min_string
  coefs$max_thresholds <- max_string
  coefs$optimum <- optimal_string
  coefs$start_time <- start_time
  coefs$timespan <- delta_t_total
  coefs$region <- region
  coefs$ssid_string <- ssid_string

  coefs <- rbind(other_coefs, coefs) %>% distinct()
  saveRDS(coefs, file = paste0("data/",climate_string,"-vocc-by-spp-sdmTMB-coefs.RDS"))

  stop("No vectors. Conditions saved to 'data/vocc-byspp-sdmTMB-coefs'.")
}
#glimpse(vocc)
#glimpse(vectors)
```

@@

biomass_change <- make_trend_data(i,
  ssid = model_ssid,
  start_time = start_time,
  skip_time = skip_time,
  end_time = end_time,
  input_cell_size = input_cell_size,
  scale_fac = scale_fac,
  delta_t_total = delta_t_total,
  delta_t_step = delta_t_step,
  indices = indices,
  variable_names = "est_exp"
)
# browser()
vectors <- vectors %>%
      mutate(
        start_cell = icell,
        vector_id = paste0(id, "_to_", tid),
        distance2 = distance*distance,
        source_end_value = e, # may need to turn into numeric vector later...
        target_value = target_values
      )

start <- biomass_change %>% 
  rename(start_density = est_exp_s, end_density = est_exp_e) %>% 
  dplyr::select(icell, x, y, start_density, end_density)

@@

end <- biomass_change %>% 
  rename(start_density = est_exp_s, 
         end_density = est_exp_e, 
         target_X = x, target_Y = y) %>% 
  dplyr::select(start_density, end_density, target_X, target_Y)

@@

source_density <- biomass_change %>% 
  rename(source_density = est_exp_s) %>% 
  dplyr::select(source_density, x, y)

@@

  # calculate percent change in biomass
  mutate(
    log_change = log(end_density/start_density),
    perc_change = end_density/start_density
    )

@@

# calculate standardized density variables
vocc_by_sp <- vocc_by_sp %>%
  mutate(
    std_source_density = (source_density - mean.density)/(sd.density*2),
    std_source_density2 = (source_density - mean.density)/(sd.density*2)^2,
    mean_source_density = mean.density, sd_density = sd.density,
    mean_change = mean.change, sd_change = sd.change
    )

@@

vocc_by_sp <- mutate(vocc_by_sp,
  (!! mean.var) := icells[1, mean.var],
  (!! sd.var) := icells[1, sd.var],
  (!! std.var) := (UQ(rlang::sym(paste0(variable_names[j], "_e"))) - icells[1, mean.var])/ (icells[1, sd.var]*2),
  (!! std.var2) := ((UQ(rlang::sym(paste0(variable_names[j], "_e"))) - icells[1, mean.var])/ icells[1, sd.var])^2
  )
}

@@

mature_plot <- lapply(variable_names, function(j) { 
ggplot(data, aes(x = UQ(rlang::sym(paste0("end_", j))), y = end_density, colour = cell_type, size = start_density, 
  group = cell_type)) +
  geom_point(alpha = 0.5) +
  # geom_smooth(method="lm", aes(x = UQ(rlang::sym(paste0("end_", j))), y = log_change
  # ), formula = y ~ x + I(x^2), inherit.aes = FALSE ) + # + I(x^3)
  # scale_size_continuous(guide = "none") +
  scale_y_continuous(trans = fourth_root_power) +
  xlab(paste("Final value of", j,"for cells")) +
  ylab(paste("Final biomass")) +
  scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
  gfplot::theme_pbs() + theme(legend.position = c(0.8, 0.8)) + 
  ggtitle(paste(survey, "mature", species))
})
print(mature_plot)
```

@@

mature_plot2 <- lapply(variable_names, function(j) { 
  ggplot(data, aes(x = distance, y = log_change, 
      group = cell_type, size = start_density, 
      colour = UQ(rlang::sym(paste0("end_", j))))) +
    geom_point(alpha = 0.5) +
    #geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
    geom_smooth(method="lm", formula = y~x+I(x^2), se = FALSE )+
    scale_size_continuous(guide = "none") +
    scale_y_continuous(trans = fourth_root_power) +
    facet_wrap(~cell_type) +#, scales = "free") +
    ylab(paste("Log change in biomass over", vectors$timespan[1], "years")) +
    xlab(paste("Distance traveled to stay within critical range of", j,"")) +
    scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", j)) + 
    # scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
    gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.2)) + 
    ggtitle(paste(survey, "mature", species))
})
mature_plot2
```

@@

mature_plot3 <- lapply(variable_names, function(j) { 
  ggplot(data, aes(x = source_density, y = log_change, 
          colour = UQ(rlang::sym(paste0("end_", j))),
    alpha = 1/(distance+ input_cell_size), #size = 1/distance,
    group = cell_type)) +
  geom_jitter(width = 0.01) +
  geom_smooth(method="lm", se = FALSE) +
  facet_wrap(~cell_type) + #, scales = "free") +
  scale_alpha_continuous(guide = "none") +
  scale_size_continuous(guide = "none") +
  scale_x_continuous(trans = sqrt) +
  scale_y_continuous(trans = sqrt) +
  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
  xlab(paste("Source density (dark points are closer to source)")) +
  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", j)) + 
  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
  ggtitle(paste(survey, "mature", species, "(mean biomass change =", round(data$mean_change[1], digits = 4), "kg/ha)"))
})
mature_plot3
```

@@

mature_plot4 <- lapply(variable_names, function(j) {  
ggplot(filter(data, distance < 10), aes(source_density, y = log_change , 
    colour = UQ(rlang::sym(paste0("end_", j))),
    alpha = 1/(distance + input_cell_size), #size = 1/distance,
    group = cell_type)) +
  geom_jitter(width = 0.01) +
  geom_smooth(method="lm", aes(group = cell_type)) +
  facet_wrap(~cell_type) +#, scales = "free") +
  scale_alpha_continuous(guide = "none") +
  scale_size_continuous(guide = "none") +
  scale_x_continuous(trans = sqrt) +
  #scale_y_continuous(trans = fourth_root_power) +
  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
  xlab(paste("Source density (dark points are closer to source)")) +
  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", j)) + 
  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
  ggtitle(paste(survey, "mature", species, "(mean biomass change =", round(data$mean_change[1], digits = 4), "kg/ha)"))
})
mature_plot4
```

@@

mature_plot5 <- lapply(variable_names, function(j) {  
  ggplot(data, aes(
      x = UQ(rlang::sym(paste0(j, ".units_per_decade"))) , 
      y = (end_density - start_density), 
      colour = cell_type, 
      size = start_density)) +
    geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
    #scale_size_continuous(guide = "none") +
    #scale_y_continuous(trans = fourth_root_power) +
    ylab(paste("kg/ha change in biomass over", vectors$timespan[1], "years")) +
    xlab(paste("Change in", j," in units per decade")) +
    scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
    gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
    ggtitle(paste(survey, "mature", species))
})
mature_plot5
```

@@

```{r}
# test <- lmerTest::lmer(log(end_density) ~ cell_type * std_source_density + cell_type * distance + 
#             std_end_climate + std_end_climate2 + 
#             (1 | vector_id), 
#  #   offset = log(start_density), 
#  #   family = Gamma(link = "inverse")
#      data = data)
# hist(data$end_density)
# hist(data$log_change)
data <- vocc_by_sp[["adult"]]

model_formula <- paste0("log_change ~ cell_type * std_source_density + cell_type * std_source_density2 + cell_type * distance ") # + cell_type * distance2

for (j in (variable_names) ){
  model_formula <- paste0("", model_formula, " + std.", j, "+ std.", j, "2"
  )
}

model_formula <- paste0("", model_formula, "+ (1 | vector_id)")

test <- lmerTest::lmer(as.formula(model_formula), REML = F, data = data)

summary(test)
# plot(test2)

library(sjPlot)
#library(sjlabelled)
#library(sjmisc)
sjPlot::plot_model(test, title = "Log biomass change")
```

```{r}
beta <- lme4::fixef(test) 
se <- arm::se.fixef(test) 
var<- names(beta)

coefs <- as.data.frame(cbind(beta, se))
coefs$ci <- coefs$se*1.96
var[1] <- "intercept"
row.names(coefs) <- NULL
coefs <-cbind(var, coefs)
coefs$model <- model_formula
coefs$species <- spp 
coefs$lifestage <- "adult"
coefs$climate <- climate 
coefs$min_thresholds <- min_thresholds
coefs$max_thresholds <- max_thresholds
coefs$optimum <- optimal_string
coefs$start_time <- start_time
coefs$timespan <- delta_t_total
coefs$region <- region
coefs$ssid_string <- ssid_string

glimpse(coefs)
rm(other_coefs)
try({other_coefs <- readRDS(paste0("data/vocc-",climate_string,"-by-spp-lmer-coefs.RDS"))})

if (exists("other_coefs")) {
coefs <- rbind(other_coefs, coefs)
}
saveRDS(coefs, file = paste0("data/vocc-",climate_string,"-by-spp-lmer-coefs.RDS"))
```

@@

test3 <- robustlmm::rlmer(as.formula(model_formula), data = data)

@@

Maybe try GAM?
```{r}
# library()
# 
# model_formula <- paste0("log_change ~ cell_type + s(std_source_density) + s(distance) ") # + cell_type * distance2
# 
# for (j in (variable_names) ){
#   model_formula <- paste0("", model_formula, " + s(std.", j, ") + "
#   )
# }
# 
# model_formula <- paste0("", model_formula, "+ (1 | vector_id)")
# 
# 
# test2 <- gamm4::....(as.formula(model_formula), REML = F, data = data)
#

@@

gbm_formula <- paste0("log_change ~ std_source_density + distance + cell_type + X + Y") # + cell_type * distance2

for (j in (variable_names) ){
  gbm_formula <- paste0(gbm_formula, " + std.", j)
}

m <- gbm::gbm(
  as.formula(gbm_formula), 
  data = data, 
  distribution = "gaussian",
  n.trees = 2000, 
  interaction.depth = 3, 
  shrinkage = 0.02
  )

@@

```{r}
plot(m,1)
plot(m,2)
plot(m,3)
plot(m,6)
if (length(variable_names) > 1 ) plot(m,7)
if (length(variable_names) > 2 ) plot(m,8)

plot(m,c(1,2))
plot(m,c(1,3))
plot(m,c(1,6))

if (length(variable_names) > 1 ) plot(m,c(1,7))

plot(m,c(2,3))
plot(m,c(2,6))
if (length(variable_names) > 1 ) plot(m,c(2,7))

plot(m,c(1,2,6))
plot(m,c(2,6,3))
plot(m,c(4,5))
```

@@

```{r}
data <- vocc_by_sp[["adult"]]
# Xmax <- quantile(data$X, 0.9995)
# Xmin <- quantile(data$X, 0.005)
# std_max <- quantile(data$log_change, 0.9995)
# 
# data <- data %>% filter ( X >= Xmin & X <= Xmax )

@@

```{r}
data <- vocc_by_sp[["adult"]]

tmb_formula <- paste0("log_change ~ cell_type * std_source_density + cell_type * distance + distance2") # + cell_type * std_source_density2

for (j in (variable_names) ){
  tmb_formula <- paste0("", tmb_formula, " + std.", j, "+ std.", j, "2"
  )
}

test <- sdmTMB(data = data,
      as.formula(tmb_formula), 
      spde = spde, # control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = TRUE)
test
```

@@

Check final climate-biomass relationship
```{r}
data <- vocc_by_sp[["imm"]]
#glimpse(data)

immature_plot <- lapply(variable_names, function(j) { 
ggplot(data, aes(x = UQ(rlang::sym(paste0("end_", j))), y = (end_density # - start_density
  ), colour = cell_type, size = start_density, 
  group = cell_type)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method="lm", aes(x = UQ(rlang::sym(paste0("end_", j))), y = end_density
  ), formula = y ~ x + I(x^2), inherit.aes = FALSE ) + # + I(x^3)
  scale_size_continuous(guide = "none") +
  scale_y_continuous(trans = fourth_root_power) +
  xlab(paste("Final value of", j,"for cells")) +
  ylab(paste("Final biomass")) +
  scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
  gfplot::theme_pbs() + theme(legend.position = c(0.8, 0.8)) + 
  ggtitle(paste(survey, "immature", species))
})
immature_plot
```

Check distance threshold of 50 looks reasonable
```{r}
data <- vocc_by_sp[["imm"]]
#data <- filter(vocc_by_sp[["imm"]],  distance < 90) # distance > 1 &
#data <- filter(vocc_by_sp[["imm"]], distance > 1 ) #  &

immature_plot2 <- lapply(variable_names, function(j) { 
  ggplot(data, aes(x = distance, y = log_change, 
      group = cell_type, size = start_density, 
      colour = UQ(rlang::sym(paste0("end_", j))))) +
    geom_point(alpha = 0.5) +
    #geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
    geom_smooth(method="lm", formula = y~x+I(x^2) )+
    scale_size_continuous(guide = "none") +
    scale_y_continuous(trans = fourth_root_power) +
    facet_wrap(~cell_type) +#, scales = "free") +
    ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
    xlab(paste("Distance traveled to stay within critical range of", j,"")) +
    scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", j)) + 
    # scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
    gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.2)) + 
    ggtitle(paste(survey, "immature", species))
})
immature_plot2
```

```{r warnings=FALSE}
data <- vocc_by_sp[["imm"]]
#data <- filter(vocc_by_sp[["imm"]], distance < 50)
#data <- filter(vocc_by_sp[["imm"]], distance > 4 & distance < 50)

immature_plot3 <- lapply(variable_names, function(j) { 
  ggplot(data, aes(x = source_density, y = log_change, 
          colour = UQ(rlang::sym(paste0("end_", j))),
    alpha = 1/(distance+ input_cell_size), #size = 1/distance,
    group = cell_type)) +
  geom_jitter(width = 0.01) +
  geom_smooth(method="lm") +
  facet_wrap(~cell_type) + #, scales = "free") +
  scale_alpha_continuous(guide = "none") +
  scale_size_continuous(guide = "none") +
  scale_x_continuous(trans = sqrt) +
  scale_y_continuous(trans = sqrt) +
  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
  xlab(paste("Source density (dark points are closer to source)")) +
  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", j)) + 
  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
  ggtitle(paste(survey, "immature", species, "(mean biomass change =", round(data$mean_change[1], digits = 4), "kg/ha)"))
})
immature_plot3
```

```{r}
data <- vocc_by_sp[["imm"]]
data <- filter(vocc_by_sp[["imm"]], distance < 50)

immature_plot4 <- lapply(variable_names, function(j) {  
ggplot(filter(data, distance < 10), aes(source_density, y = log_change , 
    colour = UQ(rlang::sym(paste0("end_", j))),
    alpha = 1/(distance + input_cell_size), #size = 1/distance,
    group = cell_type)) +
  geom_jitter(width = 0.01) +
  geom_smooth(method="lm", aes(group = cell_type)) +
  facet_wrap(~cell_type) +#, scales = "free") +
  scale_alpha_continuous(guide = "none") +
  scale_size_continuous(guide = "none") +
  scale_x_continuous(trans = sqrt) +
  #scale_y_continuous(trans = fourth_root_power) +
  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
  xlab(paste("Source density (dark points are closer to source)")) +
  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", j)) + 
  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
  ggtitle(paste(survey, "immature", species, "(mean biomass change =", round(data$mean_change[1], digits = 4), "kg/ha)"))
})
immature_plot4
```

```{r}
data <- vocc_by_sp[["imm"]]
data <- filter(vocc_by_sp[["imm"]], distance < 50)

immature_plot5 <- lapply(variable_names, function(j) {  
  ggplot(data, aes(
      x = UQ(rlang::sym(paste0(j, ".units_per_decade"))) , 
      y = (end_density - start_density), 
      colour = cell_type, 
      size = start_density)) +
    geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
    #scale_size_continuous(guide = "none") +
    #scale_y_continuous(trans = fourth_root_power) +
    ylab(paste("kg/ha change in biomass over", vectors$timespan[1], "years")) +
    xlab(paste("Change in", j," in units per decade")) +
    scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
    gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.7)) + 
    ggtitle(paste(survey, "immature", species))
})
immature_plot5
```

## Immature models

@@

```{r}
summary(im)
```
